---
title: "Comparative varianttransmissibility: Alpha vs Delta"
author: ""
date: "21/06/2021"
output: 
  rmarkdown::github_document
#rmarkdown::pdf_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning =F)
```


In these analyses we compare B.1.1.7 (alpha variant) with B.1.617.2 (delta variant) in terms of lineage sizes in Norway.
Each inferred import results in a single point in the plot below, and the lineage size is plotted as a function of estimated TMRCA



```{r remedy001, include=FALSE}
#   ____________________________________________________________________________
#   Loading required packages                                               ####

library(ggplot2)
library(ips)
library(data.table)
library(lubridate)
library(viridis)
library(grid)
library(hexbin)
library(lubridate)
library(RColorBrewer)


#   ____________________________________________________________________________
#   Setup paths - replace these to run locally                              ####

path_Main = "~/Dropbox/Covid/Alpha_vs_Delta/"

#   ____________________________________________________________________________
#   Read required data                                                      ####

dat = read.csv(paste0(path_Main,"alpha_vs_delta_2021_06_21.csv"))
dat
str(dat)
dat$TMRCA <- as.Date(dat$TMRCA)
#   ____________________________________________________________________________
#   Analysis begins here                                                    ####


```{r,echo=F,fig.width=9, fig.height=6}
g1 = ggplot(dat, aes(TMRCA, Lineage_size))+
  geom_point(aes(color = Variant), size=3)+
  theme_bw(base_size=20)+
  theme(legend.position=c(0.85,0.85))+
  scale_x_date(date_breaks="1 week")+
  scale_y_continuous(trans='log10') +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  xlab("Estimated TMRCA")+ylab("Estimated linage size")
g1
```

###### Figure 1:  Transmission lineage sizes as a function of time of introduction. “TMRCA” = time of most recent common ancestor. Transmission lineages are colored by variant.

```{r,echo=F,fig.width=9, fig.height=6}
g2 = ggplot(dat, aes(TMRCA, Lineage_size))+ 
  geom_point(aes(color = Variant), size=3)+
  theme_bw(base_size=20)+
  theme(legend.position=c(0.85,0.85))+
  scale_x_date(date_breaks="1 week",limits =c(as.Date("2021-04-01"),as.Date("2021-05-31")))+
  ylim(0,60) +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  xlab("Estimated TMRCA")+ylab("Estimated linage size")
g2
```

###### Figure 2:  Transmission lineage sizes as a function of time of introduction. “TMRCA” = time of most recent common ancestor. Transmission lineages are colored by variant. Same as above but linear scale and restricted to period of co-occurence.
