---
knit: (function(input_file, encoding) {
  out_dir <- '.';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'firstDiscoveredSouthAfrica.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning =F)
```

Phylogeographic mapping

Ancestral character estimation was performed on a dated phylogeny exported from NextStrain. Each isolate was assigned to one of two binary categories: "Norway" or "Rest of the world (RoW)". The output is presented in Fig.1.



```{r remedy001, include=FALSE}
### . . . . . . . . . . . . . . .Notes . . . . . . . . . . . . . . . . . . . . .
### . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . ..
### Sections that takes time to run are commented  out                      ####
### When a new build is available these should be un-commented, the         ####
### result saved, and recommented.                                          ####

#   ____________________________________________________________________________
#   Loading required packages                                               ####

library(ggplot2)
library(phytools)
library(ape)
library(castor)
library(ips)
library(data.table)
library(lubridate)
library(ggtree)
library(treedater)
library(viridis)
library(grid)
library(BactDating)
library(LineageHomology)
library(pbapply)
library(hexbin)


#   ____________________________________________________________________________
#   Setup paths - replace these to run locally                              ####

path_Main = "~/Dropbox/Covid/Southafrican_03_24/"
path_Results = "~/Dropbox/Covid/Southafrican_03_24/Results/"
path_Rscripts = "~/Dropbox/Covid/Rfunctions/"
path_Data =  "~/Dropbox/Covid/Southafrican_03_24/Data/"

#   ____________________________________________________________________________
#   Read required data                                                      ####
metadat = fread(paste0(path_Main,"nextstrain_groups_niph_ncov_southafrican-2021-05-10_metadata.tsv"))
metadata = metadat #Used alternate spellings some places.
tree = read.tree(paste0(path_Main,"nextstrain_groups_niph_ncov_southafrican-2021-05-10_timetree.nwk"))


#   ____________________________________________________________________________
#   Analysis begins here                                                    ####

plot.phylo(tree,show.tip.label = F,type="cladogram")
nodelabels(cex=0.5)
#Exclude the basal branch, NB: update this for new builds.
tree = extract.clade(tree, node=2916)
plot.phylo(tree,show.tip.label = F)
Locations = grepl("Norway",tree$tip.label)
Locations[Locations=="TRUE"]="Norway"
Locations[Locations=="FALSE"]="RoW"
names(Locations)=tree$tip.label
table(Locations)

#Remedy for ugly tree branches
set.seed(100)
tree = multi2di(tree)
tree$edge.length[tree$edge.length<=0]=min(tree$edge.length[tree$edge.length>0])
tree=ladderize(tree)#reorder.phylo(tree)


#   ____________________________________________________________________________
#   Ancestral state reconstruction                                          ####

#Reconstruct ancestral states using maximum likelihood.
locations_for_castor = as.numeric(as.factor(Locations))
names(locations_for_castor)=names(Locations)

castor = castor::asr_mk_model(tree,tip_states =locations_for_castor,
                              rate_model="ARD",include_ancestral_likelihoods = T)


#Setup date variables that are used in downstream function calls.
mrsd = max(metadat$`Collection Data`)
dec_date = decimal_date(max(metadat$`Collection Data`))
start_date = dec_date-max(tipHeights(tree))
most_recent_date = ymd(mrsd)
most_recent_date = decimal_date(most_recent_date)
start_time = most_recent_date-max(nodeHeights(tree))
time_end = most_recent_date-start_time
start_time = start_date
time_end = start_time+max(nodeHeights(tree))

#Find the labels of the taxa that are observed in Norway. 
Norwegian_tips = names(Locations[Locations=="Norway"])



#Lineage description from the maximum likelihood mappings obtained by sampling from node probabilities.
#multi_counts = pbreplicate(100, LineageHomology_w_uncertainty(tree, ace_nodes = castor$ancestral_likelihoods, give_tips = Norwegian_tips,ace_tips = to.matrix(Locations,seq=c("Norway","RoW")),start_time = start_date))
#save(multi_counts, file=paste0(path_Data,"count_ML_southafrica_100.Rdata"))
load(paste0(path_Data,"count_ML_southafrica_100.Rdata"))

#Lineage descriptions using the state with more than 50% probability for all nodes.
Result = LineageHomology(tree, ace_nodes = castor$ancestral_likelihoods,give_tips = Norwegian_tips,
                                 ace_tips = to.matrix(Locations,seq=c("Norway","RoW")),start_time = start_date)

```

```{r, echo=FALSE, fig.width=13, fig.height=20}
cols<-setNames( c("#E41A1C","#377EB8"),unique(Locations))
plot.phylo(tree,cex=0.3, mar=c(5.1,0.2,0.2,0.2),show.tip.label = F,edge.width=2)
LineageHomology::axisPhyloDate(1,root.time= decimal_date(mrsd)-max(tipHeights(tree)), backward = F, lwd=4)
nodelabels(pie=castor$ancestral_likelihoods,cex=0.2,piecol=cols)
tiplabels(pie=to.matrix(Locations,seq=c("Norway","RoW")),cex=0.2,piecol=cols)
#abline(v=seq(0,382,10), lty=2, col="grey")
legend("bottomleft",c("Norway", "RoW"), col = c("#E41A1C","#377EB8"),
       pch=20, cex = 3, pt.cex=3, bty="n")
# 
```
 
##### Fig. 1. B.1.351 phylogeny colored by binary geographic mappings (Norway, rest of the world [RoW]). 
 
Quantifying introductions and local transmission

Below is a visualization of the size of each transmission lineage detected in Norway.

```{r, echo=FALSE, fig.width=8, fig.height=8}
all_lineages = Result[[2]]
#Plot lineages:
library(lubridate)
library(treemap)
library(RColorBrewer)
library(ggtree)
#Treemap lineages.
atreemap = data.frame(group=paste0("G:",1:length(all_lineages)," S: ", all_lineages, " mrca: ",decimal2Date(Result$`MRCA's`)),Value=all_lineages)
treemap(atreemap,index=("group"),vSize="Value", type="index", palette="Paired")
```

##### Fig. 2. B.1.351 transmission lineage sizes in Norway. G denotes the number assigned to each lineage, S denotes the size of the lineage and TMRCA denotes the age of the lineage.

```{r,include=F}
#   ____________________________________________________________________________
#   Import export estimates                                                 ####
tip_list = Norwegian_tips
dates = decimal_date(metadat$`Collection Data`)
names(dates)=metadata$Strain
source("~/Dropbox/Github/SARS-CoV-2_phylogeo/docs/Analyses/R_functions/Relative_load_importation_local_transmission.R")
Result_with_uncertainty = Relative_load_import_multi(tree, multi_counts,dates, start_date,tip_list=Norwegian_tips)
 

#Make weekly time windows to count events. 
weeks = seq(start_time, time_end, by = 1/52)
week_time=weeks[1:(length(weeks)-1)]+1/(52*2) #This is only for returning results at midpoint of weeks.
weekly_importations = apply(Result_with_uncertainty$Import,2, FUN="mean")
weekly_local_tranmissions = apply(Result_with_uncertainty$LC,2, FUN="mean")
sum_cases = weekly_importations+weekly_local_tranmissions

#Calculate confidence intervals
imports = t(apply(Result_with_uncertainty$Import, 2,quantile,c(0.025,0.5,0.975)))
exports = t(apply(Result_with_uncertainty$LC, 2,quantile,c(0.025,0.5,0.975)))
dat=data.frame(cbind(imports,exports),dates=decimal2Date(week_time))
colnames(dat)=c("i2.5","i50","i97.5","e2.5","e50","e97.5","dates")

#Setup data for import versus local transmission based on the mean.
ggdata = data.frame(dates = as.Date(date_decimal(week_time)),
                  fraction=c(weekly_importations/sum_cases,weekly_local_tranmissions/sum_cases),
                  group=c(rep("Import",length(week_time)),
                          rep("Local transmission",length(week_time))))
```

The number of Norwegian isolates resulting from local transmission vs introductions from abroad were estimated from the tree and is summarized in Table 1.

##### Table 1. Estimated number of Norwegian infections that are caused by import and local transmission in geographically mapped phylogeny in Figure 1.

```{r, echo=F,fig.width=10, fig.height=10}
imports_temp = unlist(multi_counts[1,])[seq(1,length(multi_counts[1,])*2,by=2)]
local_temp = unlist(multi_counts[1,])[seq(2,length(multi_counts[1,])*2,by=2)]
relative_temp = imports_temp/(imports_temp+local_temp)
c_one = quantile(imports_temp,c(0.025,0.50,0.975))
c_two = quantile(local_temp,c(0.025,0.50,0.975))
c_three = quantile(relative_temp,c(0.025,0.50,0.975))
imp_loc = rbind(c_one,c_two, c_three)
rownames(imp_loc)=c("Import", "Local transmission", "Import / Total")
knitr::kable(imp_loc, booktabs = T, digits = 4)
```


```{r,echo=F, fig.width=7,fig.height=7}
g1 = ggplot(data=ggdata, aes(x=dates, y=fraction, fill=group)) + 
  geom_area(alpha=0.6 , size=1, colour="black")+
  theme_minimal(base_size=20)+scale_fill_brewer(palette="Set1")+
  scale_x_date(date_breaks="1 week")+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ylab("")+xlab("")+
  annotate("rect",xmin=as.Date(decimal2Date(time_end-1/13)),xmax=as.Date(dat$dates[length(dat$dates)]-7),ymin=-Inf,ymax=Inf,alpha=0.2)+
  theme(legend.position = c(0.55,0.15), legend.title=element_blank())
g1
```

###### Fig. 3. Relative quantification of infection source (new introduction / import vs local transmission).


```{r,include=F}
g1 = ggplot(dat, aes(x=dates,y=i50))+geom_line(col="Red")+theme_minimal(base_size=20)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  ylim(0,5)+
  geom_ribbon(aes(ymin=i2.5, ymax=i97.5,alpha=0.2), fill="lightblue")+
  scale_x_date(date_breaks="1 week",limits =c(as.Date("2020-11-30"), as.Date("2021-03-28")),expand=c(0,0))+
  ylab("Import")+
  theme(legend.position = "none")+annotate("rect",xmin=as.Date(decimal2Date(time_end-1/13)),xmax=as.Date(dat$dates[length(dat$dates)])-7,ymin=-Inf,ymax=Inf,alpha=0.2)
g1

g2 = ggplot(dat, aes(x=dates,y=e50))+geom_line(col="Darkgreen")+theme_minimal(base_size=20)+ylim(0,80)+
  scale_x_date(date_breaks="1 week",limits =c(as.Date("2020-11-30"), as.Date("2021-03-28")), expand=c(0,0))+
  geom_ribbon(aes(ymin=e2.5, ymax=e97.5,alpha=0.2), fill="lightblue")+
  theme(axis.text.x=element_text(angle=60, hjust=1))+scale_fill_continuous("Grey")+ylab("Local transmission")+xlab("")+
  theme(legend.position = "none")+annotate("rect",xmin=as.Date(decimal2Date(time_end-1/13)),xmax=as.Date(dat$dates[length(dat$dates)])-7,ymin=-Inf,ymax=Inf,alpha=0.2)
g2
```

```{r,echo=F,fig.width=10, fig.height=10}
grid.draw(rbind(ggplotGrob(g1), ggplotGrob(g2), size = "last"))
```

###### Fig. 4. Absolute quantification of imports and local transmission events. 

```{r,echo=F,fig.width=9, fig.height=6}
c2 = data.frame(decimal2Date(cbind(Result$`MRCA's`)),Result$Lineage_sizes)
colnames(c2)=c("MRCA","Lineagesize")
g1 = ggplot(c2, aes(MRCA, Lineagesize))+
  geom_hex()+
  theme_bw(base_size=20)+
  theme(legend.position=c(0.85,0.85))+
  scale_fill_viridis()+
  scale_x_date(date_breaks="1 week")+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  xlab("Estimated TMRCA")+ylab("Estimated linage size")+
  annotate("rect",xmin=as.Date(decimal2Date(time_end-1/12)),xmax=as.Date(dat$dates[length(dat$dates)]),ymin=-Inf,ymax=Inf,alpha=0.2)
g1

```

###### Fig. 5.  Transmission lineage sizes as a function of time of introduction. “TMRCA” = time of most recent common ancestor.
